### 2020-06-19 submission/manuscript.tex
Render with Ada's comments

To address this need, we noted the similarity between the feeling of
unease in analyzing the data and the struggles scientists face with
engaging the voluminous scientific literature. A common strategy for
keeping up with the literature is by participating in a journal club.
Regardless of the discipline, a Journal Club involves group discussion
of a pre-selected paper. These range from informal discussions to
PowerPoint presentations and course credit. In addition to building upon
material from traditional coursework and staying current on the
literature, journal clubs help strengthen skills in critical thinking,
communication, and integrating the literature {[}11{]}. Since most
journal clubs occur on a regular schedule, they are effective because
they incorporate repeated deliberate practices. With this model in mind,
over the past four years we have experimented with creating a Code Club
model with the goal of improving reproducible data analysis skills in a
community environment.

@@

From these experiences, we began a collective conversation to improve
our Code Club model. We have identified two successful approaches. The
first is a more constructive version of a group code review. The
presenter clearly states the problem they want to solve, breaks the
participants into smaller groups, and then asks each group to solve the
problem, or a portion of it. For example, someone may have an R script
with a repeating chunk of code. The challenge for the session would be
to convert the code chunk into a function to be called throughout the
script to make it ``DRY'' (i.e.~Don't Repeat Yourself {[}12{]}). The
presenter leaves the session with several partial or working solutions
to their problem and the importance of writing DRY code is reinforced.
The second approach is a tutorial. The presenter introduces a new
package or technique and assigns an activity to practice the new
approach. For example, at one Code Club participants were given raw data
and a finished plot. Paired participants were tasked with generating the
plot from the data using either R syntax from the base language or the
ggplot2 package. For this exercise, base R users had to use ggplot2 and
vice versa. In either approach, the Code Club ends with a report back to
the larger group describing the approach each pair took. We generally
find that preparing a Code Club session takes as much effort as
preparing to present at a journal club. Our Code Clubs typically have 7
to 10 participants, but the inherent ``think-pair-share'' approach
should allow it to be scaled to larger groups {[}13{]}.

We continue to experiment with approaches for running Code Club. For
example, during the period of the Covid-19 pandemic in 2020 when many
research labs were shuttered, PDS started posting Code Club materials
(\url{https://www.riffomonas.org/code_club/}) that individuals or
research groups can use to practice their coding skills. During this
time, we continued our group's Code Club sessions by making use of the
breakout room and screen sharing features in Zoom. Regardless of our
approach to leading a Code Club, we have learned that it is critical for
the presenter to clearly articulate their goals and facilitate
participant engagement. Although some Code Club sessions may be more
experimental than others, on the whole they are a critical tool to train
bench scientists in reproducible data analysis practices. We have
provided some examples of successful topics in Table 1. We have
summarized the results of our own experimentation as Code Club
presenters and participants into Ten Rules. The first 3 rules apply to
all participants, Rules 4 through 8 are targeted to presenters, and the
last two for non-presenters. As the rules describe, this model can
easily be adapted to groups of non-biologists and to groups with
stronger background in computer science.

@@

As suggested by Rule 7, a significant challenge to presenting at Code
Club is selecting topics and activities that appeal to a critical mass
of the participants. This is particularly difficult if participants have
a wide range of coding experience, which can follow a turnover in group
membership. Beginners will benefit from sessions that cover fundamental
concepts and functions. The benefits experienced by more experienced
participants include the improved understanding of concepts that comes
through teaching the concepts and breaking down problems into simpler
elements. Furthermore, instead of focusing on core functions, the group
could balance Code Club sessions that cover basics with those that cover
methods and packages that are new to everyone in the group. We have also
identified several strategies to overcome the challenges presented by
participants at varying skill levels. Central to the Code Club format is
the use of paired-programming {[}15{]}. Instead of letting participants
form their own pairs, the presenter can select pairs of participants
with either similar or differing skill levels, depending on the
presenter's goals. Partnerships between those with similar skill levels
requires the presenter to design appropriate activities for each skill
level. We have found that commenting code is a good skill for beginners
since it forces them to dissect and understand a code chunk
line-by-line. It also reinforces the value of commenting as they develop
their skills and independence. An advantage of forming partnership
between people with disparate skill levels is that it is more likely for
groups to provide the presenter with a diverse range of methods that
achieve the same result. This approach to pairing also helps to graft
new members that have emerging programming skills. Regardless of how
partners are selected, consider asking the pairs to identify a navigator
and a driver {[}16{]}. The driver types at the computer while the
navigator tells them what to type, thus ensuring participation of both
partners. Midway through the activity, the presenter can have the
partners switch roles. Intentionally forming pairs can also engineer
group interactions by avoiding potentially disruptive partnerships or
pairing reliable role models with new group members.

>>>>

To address this need, we noted the similarity between the feeling of
unease in analyzing the data and the struggles scientists face with
engaging the voluminous scientific literature. A common strategy for
keeping up with the literature is by participating in a journal club.
Regardless of the discipline, a Journal Club involves group discussion
of a pre-selected paper that range from informal discussions to
PowerPoint presentations and course credit. In addition to building upon
material from traditional coursework and staying current on the
literature, journal clubs help strengthen skills in critical thinking,
communication, and integrating the literature {[}11{]}. Since most
journal clubs occur on a regular schedule, they are effective by virtue
of repeated practice. With this model in mind, over the past four years
we have experimented with creating a Code Club model with the goal of
improving reproducible data analysis skills in a community environment.

@@

From these experiences, we began a collective conversation to improve
our Code Club model. We have identified two successful approaches. The
first is a more constructive version of a group code review. The
presenter clearly states the problem they want to solve, breaks the
participants into smaller groups, and then asks each group to solve the
problem, or a portion of it. For example, someone may have an R script
with a repeating chunk of code. The challenge for the session would be
to convert the code chunk into a function to be called throughout the
script to make it ``DRY'' (i.e.~Don't Repeat Yourself {[}12{]}). The
presenter leaves the session with several partial or working solutions
to their problem and the importance of writing DRY code is reinforced.
The second approach is a tutorial. The presenter introduces a new
package or technique and assigns an activity to practice the new
approach. For example, at one Code Club participants were given raw data
and a finished plot. Paired participants were tasked with generating the
plot from the data using R syntax from either the base language or the
ggplot2 package. For this exercise, base R users had to use ggplot2 and
vice versa. In either approach, the Code Club ends with a report back to
the larger group describing the approach each pair took. We generally
find that preparing a Code Club session takes similar effort to
preparing a journal club presentation. Our Code Clubs typically have 7
to 10 participants, but the inherent ``think-pair-share'' approach
should allow it to be scaled to larger groups {[}13{]}.

We continue to experiment with approaches for running Code Club. For
example, during the early period of the Covid-19 pandemic in 2020 when
many research labs were shuttered, PDS started posting Code Club
materials (\url{https://www.riffomonas.org/code_club/}) that individuals
or research groups can use to practice their coding skills. During this
time, we continued our group's Code Club sessions using the breakout
room and screen sharing features in Zoom. Regardless of our approach to
leading a Code Club, we have learned that it is critical for the
presenter to clearly articulate their goals and facilitate participant
engagement. Although some Code Club sessions may be more experimental
than others, on the whole they are a critical tool to train bench
scientists in reproducible data analysis practices. We have provided
some examples of successful topics in Table 1. We have summarized the
results of our own experimentation as Code Club presenters and
participants into Ten Rules. The first 3 rules apply to all
participants, Rules 4 through 8 are targeted to presenters, and the last
two for non-presenters. As the rules describe, this model can easily be
adapted to groups of non-biologists or those with stronger backgrounds
in computer science.

@@

As suggested by Rule 7, a significant challenge to presenting at Code
Club is selecting topics and activities that appeal to a critical mass
of the participants. This is particularly difficult if participants have
a wide range of coding experience, which can follow a turnover in group
membership. Beginners will benefit from sessions that cover fundamental
concepts and functions. The benefits for more experienced participants
include the improved understanding of concepts by teaching and breaking
down problems into simpler elements. Furthermore, instead of focusing on
core functions, the group could balance Code Club sessions that cover
basics with those that introduce new methods and packages to the group.
We have also identified several strategies to overcome the challenges
presented by participants at varying skill levels. Central to the Code
Club format is the use of paired-programming {[}15{]}. Instead of
letting participants form their own pairs, the presenter can select
pairs of participants with either similar or differing skill levels,
depending on the presenter's goals. Partnerships between those with
similar skill levels requires the presenter to design appropriate
activities for each skill level. We have found that commenting code is a
good skill for beginners since it forces them to dissect and understand
a code chunk line-by-line. It also reinforces the value of commenting as
they develop their skills and independence. An advantage of forming
partnership between people with disparate skill levels is that it is
more likely for groups to provide the presenter with a diverse range of
methods that achieve the same result. This approach to pairing also
helps to graft new members that have emerging programming skills.
Regardless of how partners are selected, consider asking the pairs to
identify a navigator and a driver {[}16{]}. The driver types at the
computer while the navigator tells them what to type, thus ensuring
participation of both partners. Midway through the activity, the
presenter can have the partners switch roles. Intentionally forming
pairs can also engineer group interactions by avoiding potentially
disruptive partnerships or pairing reliable role models with new group
members.
